---
title: "Dashboard_leaflet"
output: html_document
date: "2025-01-23"
---
hi
```{r}
install.packages("leaflet")
library(leaflet)

leaflet() %>%  # Start the leaflet map
  addTiles() %>%  # Add default OpenStreetMap tiles
  addMarkers(lng = -122.4194, lat = 37.7749, popup = "San Francisco")
```

```{r}
install.packages("maps")
library(maps)

# Get map data for California
california_map <- map("state", region = "california", plot = FALSE, fill = TRUE)

# Create a leaflet map
leaflet() %>%
  addTiles() %>%  # Add default OpenStreetMap tiles
  addPolygons(data = california_map,
              color = "blue",
              weight = 1,
              fillOpacity = 0.5,
              label = "California") %>%
  setView(lng = -119.4179, lat = 36.7783, zoom = 6)  # Center map on California




```

```{r}
# Load necessary libraries
library(leaflet)
library(maps)
library(sf)

# Get California map data
california_map <- map("state", region = "california", plot = FALSE, fill = TRUE)

# Convert map data to sf format
california_sf <- st_as_sf(map("state", region = "california", plot = FALSE, fill = TRUE, 
                              col = "transparent"))

# Create a leaflet map with California outlined in a thick black line
leaflet() %>%
  addTiles() %>%  # Add default OpenStreetMap tiles
  addPolylines(data = california_sf,
               color = "black",     # Border color
               weight = 3) %>%      # Thickness of the border
  setView(lng = -119.4179, lat = 36.7783, zoom = 6)  # Center map on California

```

```{r}
# Create a leaflet map with white background and borders
leaflet() %>%
  addProviderTiles("CartoDB.PositronNoLabels") %>%  # White background without labels
  addPolygons(data = california,
              color = "black",      # Border color for state
              weight = 3,          # Border thickness for state
              fillOpacity = 0,     # Transparent interior for state
              label = ~NAME) %>%   # Add a label for California
  addPolygons(data = california_counties,
              color = "grey",      # Border color for counties
              weight = 1,          # Border thickness for counties
              fillOpacity = 0,     # Transparent interior for counties
              label = ~NAME) %>%   # Add a label for counties
  setView(lng = -119.4179, lat = 36.7783, zoom = 6)%>%  # Center the map on California
  setMaxBounds(
    lng1 = california_bbox["xmin"],  # West longitude limit
    lat1 = california_bbox["ymin"],  # South latitude limit
    lng2 = california_bbox["xmax"],  # East longitude limit
    lat2 = california_bbox["ymax"]   # North latitude limit
  )

```

```{r}
retailers <- d[!is.na(d$lat) & !is.na(d$lon), ]
```

```{r}
leaflet(data = retailers) %>%
  addTiles() %>%
  addMarkers(
    ~lon, ~lat,
    popup = ~paste(
      "<b>Name:</b>", name, "<br>",
      "<b>Status:</b>", business_status, "<br>",
      "<b>Rating:</b>", rating, "(", user_ratings_total, "ratings)<br>",
      "<b>Address:</b>", vicinity, "<br>",
      "<b>Website:</b>", paste0('<a href="', website, '" target="_blank">', website, "</a>")
    )
  )
```

\# Create the map with solid blue circle markers and detailed popups

leaflet(data = retailers) %\>%

addTiles() %\>%

addCircleMarkers(

\~lon, \~lat, \# Longitude and Latitude for markers

radius = 6, \# Size of the circle

fillColor = "purple", \# Fill color of the circle

fillOpacity = 1, \# Full opacity to make them solid

stroke = FALSE, \# No border around the circles

popup = \~paste(

\# HTML content for the popup

"\<b\>Name:\</b\>", name, "\<br\>",

"\<b\>Status:\</b\>", business_status, "\<br\>",

"\<b\>Rating:\</b\>", rating, "(", user_ratings_total, "ratings)\<br\>",

"\<b\>Address:\</b\>", vicinity, "\<br\>",

"\<b\>Website:\</b\>", ifelse(

!is.na(website),

paste0('\<a href="', website, '" target="\_blank"\>', website, "\</a\>"),

"No website available"

)

)

)

```{r}

# Create the map with solid blue circle markers and detailed popups
leaflet(data = retailers) %>%
  addTiles() %>%
  addCircleMarkers(
    ~lon, ~lat,               # Longitude and Latitude for markers
    radius = 6,               # Size of the circle
    fillColor = "purple",       # Fill color of the circle
    fillOpacity = 1,          # Full opacity to make them solid
    stroke = FALSE,           # No border around the circles
    popup = ~paste(
      # HTML content for the popup
      "<b>Name:</b>", name, "<br>",
      "<b>Status:</b>", business_status, "<br>",
      "<b>Rating:</b>", rating, "(", user_ratings_total, "ratings)<br>",
      "<b>Address:</b>", vicinity, "<br>",
      "<b>Website:</b>", ifelse(
        !is.na(website), 
        paste0('<a href="', website, '" target="_blank">', website, "</a>"), 
        "No website available"
      )
    )
  )



```

```{r}
# Add a new column to the dataset for marker color
retailers$marker_color <- ifelse(!is.na(retailers$website), "red", "blue")

# Create the map with conditional marker colors
leaflet(data = retailers) %>%
  addTiles() %>%
  addCircleMarkers(
    ~lon, ~lat,               # Longitude and Latitude for markers
    radius = 6,               # Size of the circle
    fillColor = ~marker_color, # Conditional fill color (red or blue)
    fillOpacity = 1,          # Full opacity to make them solid
    stroke = FALSE,           # No border around the circles
    popup = ~paste(
      # HTML content for the popup
      "<b>Name:</b>", name, "<br>",
      "<b>Status:</b>", business_status, "<br>",
      "<b>Rating:</b>", rating, "(", user_ratings_total, "ratings)<br>",
      "<b>Address:</b>", vicinity, "<br>",
      "<b>Website:</b>", ifelse(
        !is.na(website), 
        paste0('<a href="', website, '" target="_blank">', website, "</a>"), 
        "No website available"
      )
    )
  )


```

```{r}

library(dplyr)

# Filter the dataset to remove rows where all relevant columns are NA
retailers_clean <- retailers %>%
  filter(
    !is.na(name) | 
    !is.na(business_status) | 
    !is.na(rating) | 
    !is.na(vicinity) | 
    !is.na(website)
  )

# Create a temporary version of the dataset, flattening only the necessary columns
temp_data <- retailers_clean %>%
  mutate(
    name = unlist(name),
    business_status = unlist(business_status),
    rating = unlist(rating),
    vicinity = unlist(vicinity),
    website = unlist(website),
    marker_color = ifelse(!is.na(website), "red", "blue")
  )

# Create the map with the temporary dataset
leaflet(data = temp_data) %>%
  addTiles() %>%
  addCircleMarkers(
    ~lon, ~lat,               # Longitude and Latitude for markers
    radius = 6,               # Size of the circle
    fillColor = ~marker_color, # Conditional fill color (red or blue)
    fillOpacity = 1,          # Full opacity to make them solid
    stroke = FALSE,           # No border around the circles
    popup = ~paste(
      # HTML content for the popup
      "<b>Name:</b>", ifelse(!is.na(name), name, "No name available"), "<br>",
      "<b>Status:</b>", ifelse(!is.na(business_status), business_status, "No status available"), "<br>",
      "<b>Rating:</b>", ifelse(!is.na(rating), paste(rating, "(", user_ratings_total, "ratings)"), "No ratings available"), "<br>",
      "<b>Address:</b>", ifelse(!is.na(vicinity), vicinity, "No address available"), "<br>",
      "<b>Website:</b>", ifelse(
        !is.na(website), 
        paste0('<a href="', website, '" target="_blank">', website, "</a>"), 
        "No website available"
      )
    )
  )


```

```{r}
library(tigris)
library(leaflet)
library(sf)  # For spatial data manipulation

# Options for tigris to use sf objects
options(tigris_use_cache = TRUE)
options(tigris_class = "sf")

# Load California counties using tigris
california_counties <- counties(state = "CA", cb = TRUE)

# Filter for San Diego County
san_diego_county <- california_counties %>%
  filter(NAME == "San Diego")

# Create the leaflet map for San Diego County
leaflet() %>%
  addProviderTiles("CartoDB.PositronNoLabels") %>%  # White background without labels
  addPolygons(
    data = san_diego_county,
    color = "black",      # Border color for San Diego County
    weight = 3,           # Border thickness
    fillOpacity = 0,      # Transparent interior
    label = ~NAME         # Add a label for San Diego County
  ) %>%
  setView(
    lng = -117.1611,      # Center the map on San Diego
    lat = 32.7157,
    zoom = 9              # Adjust zoom level for San Diego County
  )

```

```{r}
# Load necessary libraries
library(tigris)
library(leaflet)
library(sf)        # For spatial data manipulation
library(dplyr)     # For data manipulation

# Options for tigris to use sf objects
options(tigris_use_cache = TRUE)
options(tigris_class = "sf")

# Step 1: Load California counties and filter for San Diego County
california_counties <- counties(state = "CA", cb = TRUE)

# Filter for San Diego County
san_diego_county <- california_counties %>%
  filter(NAME == "San Diego")

# Step 2: Prepare retailers' data
# Filter the dataset to remove rows where all relevant columns are NA
retailers_clean <- retailers %>%
  filter(
    !is.na(name) | 
    !is.na(business_status) | 
    !is.na(rating) | 
    !is.na(vicinity) | 
    !is.na(website)
  )

# Create a temporary version of the dataset, flattening only the necessary columns
temp_data <- retailers_clean %>%
  mutate(
    name = unlist(name),
    business_status = unlist(business_status),
    rating = unlist(rating),
    vicinity = unlist(vicinity),
    website = unlist(website),
    marker_color = ifelse(!is.na(website), "red", "blue")
  )

# Step 3: Combine the San Diego County boundary with retailer markers on a leaflet map
leaflet() %>%
  # Add a white background without labels
  addProviderTiles("CartoDB.PositronNoLabels") %>%
  
  # Add the San Diego County boundary
  addPolygons(
    data = san_diego_county,
    color = "black",      # Border color for San Diego County
    weight = 3,           # Border thickness
    fillOpacity = 0,      # Transparent interior
    label = NULL         # Add a label for San Diego County
  ) %>%
  
  # Add retailer markers
  addCircleMarkers(
    data = temp_data,
    ~lon, ~lat,               # Longitude and Latitude for markers
    radius = 6,               # Size of the circle
    fillColor = ~marker_color, # Conditional fill color (red or blue)
    fillOpacity = 1,          # Full opacity to make them solid
    stroke = FALSE,           # No border around the circles
    popup = ~paste(
      # HTML content for the popup
      "<b>Name:</b>", ifelse(!is.na(name), name, "No name available"), "<br>",
      "<b>Status:</b>", ifelse(!is.na(business_status), business_status, "No status available"), "<br>",
      "<b>Rating:</b>", ifelse(!is.na(rating), paste(rating, "(", user_ratings_total, "ratings)"), "No ratings available"), "<br>",
      "<b>Address:</b>", ifelse(!is.na(vicinity), vicinity, "No address available"), "<br>",
      "<b>Website:</b>", ifelse(
        !is.na(website), 
        paste0('<a href="', website, '" target="_blank">', website, "</a>"), 
        "No website available"
      )
    )
  ) %>%
  
  # Center the map on San Diego
  setView(
    lng = -117.1611,      # Center the map on San Diego
    lat = 32.7157,
    zoom = 9              # Adjust zoom level for San Diego County
  )


```

```{r}
library(tigris)
library(leaflet)
library(sf)        # For spatial data manipulation
library(dplyr)     # For data manipulation

# Options for tigris to use sf objects
options(tigris_use_cache = TRUE)
options(tigris_class = "sf")

# Step 1: Load California counties and filter for San Diego County
california_counties <- counties(state = "CA", cb = TRUE)

# Filter for San Diego County
san_diego_county <- california_counties %>%
  filter(NAME == "San Diego")

# **Fix CRS issue: Transform to WGS84**
san_diego_county <- st_transform(san_diego_county, crs = 4326)

# Step 2: Prepare retailers' data
# Filter the dataset to remove rows where all relevant columns are NA
retailers_clean <- retailers %>%
  filter(
    !is.na(name) | 
    !is.na(business_status) | 
    !is.na(rating) | 
    !is.na(vicinity) | 
    !is.na(website)
  )

# Create a temporary version of the dataset, flattening only the necessary columns
temp_data <- retailers_clean %>%
  mutate(
    name = unlist(name),
    business_status = unlist(business_status),
    rating = unlist(rating),
    vicinity = unlist(vicinity),
    website = unlist(website),
    marker_color = ifelse(!is.na(website), "red", "blue")
  )

# Step 3: Combine the San Diego County boundary with retailer markers on a leaflet map
leaflet() %>%
  # Add a white background without labels
  addProviderTiles("CartoDB.PositronNoLabels") %>%
  
  # Add the San Diego County boundary
  addPolygons(
    data = san_diego_county,
    color = "black",      # Border color for San Diego County
    weight = 3,           # Border thickness
    fillOpacity = 0,      # Transparent interior
    label = NULL         # Add a label for San Diego County
  ) %>%
  
  # Add retailer markers
  addCircleMarkers(
    data = temp_data,
    ~lon, ~lat,               # Longitude and Latitude for markers
    radius = 6,               # Size of the circle
    fillColor = ~marker_color, # Conditional fill color (red or blue)
    fillOpacity = 1,          # Full opacity to make them solid
    stroke = FALSE,           # No border around the circles
    popup = ~paste(
      # HTML content for the popup
      "<b>Name:</b>", ifelse(!is.na(name), name, "No name available"), "<br>",
      "<b>Status:</b>", ifelse(!is.na(business_status), business_status, "No status available"), "<br>",
      "<b>Rating:</b>", ifelse(!is.na(rating), paste(rating, "(", user_ratings_total, "ratings)"), "No ratings available"), "<br>",
      "<b>Address:</b>", ifelse(!is.na(vicinity), vicinity, "No address available"), "<br>",
      "<b>Website:</b>", ifelse(
        !is.na(website), 
        paste0('<a href="', website, '" target="_blank">', website, "</a>"), 
        "No website available"
      )
    )
  ) %>%
  
  # Center the map on San Diego
  setView(
    lng = -117.1611,      # Center the map on San Diego
    lat = 32.7157,
    zoom = 9              # Adjust zoom level for San Diego County
  )

```

```{r}
# Replace with the correct file path to your shapefile
shapefile_path <- "/Users/prosperityland/Downloads/Merged_Census_SDI_data/Merged_Census_SDI.shp"
shape_data <- st_read(shapefile_path)

```

```{r}
# Transform the CRS to WGS84
shape_data_wgs84 <- st_transform(shape_data, crs = 4326)

# Verify the new CRS
st_crs(shape_data_wgs84)

library(leaflet)

leaflet(data = shape_data_wgs84) %>%
  addTiles() %>%  # Add OpenStreetMap tiles
  addPolygons(
    color = "blue",  # Polygon boundary color
    weight = 1,      # Polygon boundary thickness
    fillColor = "orange",  # Fill color for polygons
    fillOpacity = 0.5      # Transparency of fill color
  )

```

```{r}

colnames(d)

```

```{r}
head(d)
```
